---
output: github_document
---

```{r, echo = FALSE, message=F}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  dev = "png",
  dpi = 500,
  fig.align = "center",
  knitr::opts_chunk$set(comment = NA)
  )
library(ggplot2)
library(BGGM)
```

#  BGGM: Bayesian Gaussian Graphical Models <img src="joss_paper/hex.pn" align="right" alt="" width="120" />



[![CRAN Version](http://www.r-pkg.org/badges/version/BGGM)](https://cran.r-project.org/package=BGGM)
[![Downloads](https://cranlogs.r-pkg.org/badges/BGGM)](https://cran.r-project.org/package=BGGM)
[![Build Status](https://travis-ci.org/donaldRwilliams/BGGM.svg?branch=master)](https://travis-ci.org/donaldRwilliams/BGGM)



The `R` package **BGGM** provides tools for making Bayesian inference in 
Gaussian graphical models (GGM). The methods are organized around two general 
approaches for Bayesian inference: (1) estimation and (2) hypothesis 
testing. The key distinction is that the former focuses on either 
the posterior or posterior predictive distribution [@Gelman1996a; see section 5 in @rubin1984bayesianly] , whereas the latter focuses on model comparison with the Bayes factor [@Jeffreys1961; @Kass1995].

## Installation

To install the latest release version (1.0.0) from CRAN use
```{r gh-installation, eval = FALSE}	
install.packages("BGGM")	
```

The current developmental version can be installed with	

```{r, eval = FALSE}	
if (!requireNamespace("remotes")) {	
  install.packages("remotes")	
}	
remotes::install_github("donaldRwilliams/BGGM")
```

Note that the developmental version is recommended, due to the algorithms being written in `c++` and 
support for all data types. The developmental version is essentially
**BGGM** version 2.0.0.



## Overview
The methods in **BGGM** build upon existing algorithms that are well-known in the literature.
The central contribution of **BGGM** is to extend those approaches:

1.  Bayesian estimation with the novel matrix-F prior distribution [@Mulder2018]
  
    + [Estimation](#bayesian-estimation) [@Williams2019]

2. Bayesian hypothesis testing with the matrix-F prior distribution [@Williams2019_bf]

    + [Exploratory hypothesis testing](#Exploratory)
  
    + [Confirmatory hypothesis testing](#Confirmatory)
    
3. Comparing Gaussian graphical models [@Williams2019; @williams2020comparing]
    
    + [Partial correlation differences](#partial-correlation-differences) 
    
    + [Posterior predictive check](#posterior-predictive-check)
    
    + [Exploratory hypothesis testing](#exploratory-groups) 
    
    + [Confirmatory hypothesis testing](#confirmatory-groups)

4. Extending inference beyond the conditional (in)dependence structure [@Williams2019]

    +  [Predictability](#Predictability) 
    
    +  [Posterior uncertainty intervals](#partial-correlation-differences) for the 
       partial correlations
       
    +  [Custom Network Statistics](#custom-network-statistics)
    
    
The computationally intensive tasks are written in `c++` via the `R` package **Rcpp** [@eddelbuettel2011rcpp] and the `c++` library **Armadillo** [@sanderson2016armadillo]. The Bayes factors are computed with the `R` package **BFpack** [@mulder2019bfpack]. Furthermore, there are [plotting](#example-network-plot) functions
for each method, control variables can be included in the model (e.g., `~ gender`), 
and there is support for missing values (see `bggm_missing`).

## Supported Data Types

* **Continuous**: The continuous method was described in  @Williams2019. Note that 
                  this is based on the customary [Wishart distribution](https://en.wikipedia.org/wiki/Wishart_distribution).

* **Binary**: The binary method builds directly upon @talhouk2012efficient
  that, in turn, built upon the approaches of @lawrence2008bayesian and
  @webb2008bayesian (to name a few).
  
* **Ordinal**: The ordinal methods require sampling thresholds. There are two approach 
   included in **BGGM**. The customary approach described in @albert1993bayesian 
   (the default) and the 'Cowles' algorithm described in @cowles1996accelerating.
   
* **Mixed**: The mixed data (a combination of discrete and continuous) method was introduced
 in @hoff2007extending. This is a semi-parametric copula model
 (i.e., a copula GGM) based on the ranked likelihood. Note that this can be used for 
 *only* ordinal data (not restricted to "mixed" data).
