---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# BGGM
This package is described in Williams and Mulder (2019) and Williams (2018). The methods are seperated into two Bayesian approaches for inference: hypothesis testing and estimation. The former is described in Williams and Mulder (2018a), and allows for testing for the presence of edges with the Bayes factor. One-sided hypothesis testing is also possible. These methods can also provide evidence for the null hypothesis. There are extensions for confirmatory hypothesis testing in GGMs, that can include inequality or equality contraints on the partial correlation. 

The estimation based method are described in Williams (2018). The methods offer advantages compared to classical methods, in that a 
measure of uncertainty is provided for all parameters. For example, each node has a distribution for the variance explained. Measures of out-of-sample performance are also available. The model is selected with credible interval exclusion of zero.

Williams, D. R. (2018, September 20). Bayesian Inference for Gaussian Graphical Models: Structure Learning, Explanation, and Prediction. [pre-print](https://doi.org/10.31234/osf.io/x8dpr)

Williams, D. R., & Mulder, J. (2019, January 14). Bayesian Hypothesis Testing for Gaussian Graphical Models:Conditional Independence and Order Constraints. [pre-print](https://doi.org/10.31234/osf.io/ypxd8)

## Installation

You can install BGGM from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("donaldRwilliams/BGGM")
```

## Example
###  Hypothesis Testing
These methods allow for gaining evidence for both conditional depedence ($\rho \neq 0$) and independence ($\rho = 0$). Note that GGMs are often thought to characterize conditional independece structures, but evidence for the null hypothesis of no effect is not typically assessed.


```{r example}
## basic example code
```
