---
title: "Predictability with Bayesian Variance Explained"
author: "Donald R. Williams"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
  toc: yes
vignette: >
  %\VignetteIndexEntry{Predictability with Bayesian Variance Explained}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
  \usepackage{amsmath}
params:
  EVAL: !r identical(Sys.getenv("NOT_CRAN"), "true")
---
  
```{r, SETTINGS-knitr, include=FALSE}
stopifnot(require(knitr))
options(width = 90)
opts_chunk$set(
  comment = NA,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  eval = if (isTRUE(exists("params"))) params$EVAL else FALSE,
  dev = "png",
  dpi = 150,
  fig.asp = 0.8,
  fig.width = 7,
  out.width = "60%",
  fig.align = "center"
)
```


# Predictability
This tutorial provides the *work flow* in **BGGM** for computing predictability in Gaussian graphical models. I will use data from a resilience questionnaire.

```{r}
# packages
library(BGGM)
library(ggplot2)

# resilence data
# remove gender variable
dat <- subset(rsa, select = - gender)
```

Note that most of the data sets in **BGGM** include categorical variables such as gender. Thus it is important to check the documentation to ensure that only the relevant variables are included in the analysis.

## In-Sample Predictability
### Estimate the Network
The first step is to estimate the network. This provides the necessary ingredients 
for computing network predictability.

```{r}
# fit model
fit <- estimate(dat, iter = 1000)
```

### Compute Bayesian Variance Explained
Next the object `fit` is used to compute Bayesian variance explained.
```{r}
# predict
pred <- predict(fit,  cred = 0.95)
```

Note `cred` serves as both the decision rule for selecting edges to include in  the network and summarizing the posterior distributions.

### Summarize Node Predictability
```{r}
# summary
pred
```

### Plotting Predictability
Most functions in **BGGM** have plots associated with them. In most cases, this 
has been simplified by calling `plot`.

```{r}
# plot 
plot(pred)
```

This is not the most attractive plot, which is by design. The returned object is a `ggplot` which can then be further customized. Here is an example with the items not colored according to their respective domain.

```{r, out.width= '100%'}
plot(pred) + 
  xlab("Item") +
  geom_point(aes(color = BGGM:::rsa_labels), 
             size = 4) +
   geom_point(size = 3, 
              color = "white") +
  scale_color_brewer(name = "Domain", 
                     palette = "Set1") +
  ggtitle("Predictability") +
  coord_cartesian() +
  theme(axis.text.x = element_text(angle = 60, 
                                   hjust = .5, 
                                   vjust = .5))
```


## Out-of-Sample Predictability
It is also possible to assess out-of-sample predictability. By this I am referring to predicting new data (test data) that was not used to select the initial network (training data).

```{r}
# training data
train_dat <- dat[1:600,]

# testing data
test_dat <- dat[601:675,]

# fit model
fit <- estimate(train_dat, iter = 1000)
```

This is the same as in-sample prediction. But now we included the test data in the predict call.
```{r}
pred <- predict(fit, 
                test_data = test_dat, 
                cred = 0.95)

pred
``` 

The results can then be plotted.
```{r}
plot(pred) + 
  ggtitle("Out-of-Sample Predictability")
```

## Custom plots
There are a variety of cool plots that can be made. One such example is the so called ridgeline plot. This can be made by accessing the posterior samples in the `pred` object.

```{r, fig.height=80}
library(ggridges)
library(ggplot2)
dat_res <- reshape2::melt(pred$post_samples)

dat_res$L1 <- factor(dat_res$L1, 
                     labels = order(pred$summary_error$post_mean, decreasing = F), 
                     levels = colnames(dat)[order(pred$summary_error$post_mean, decreasing = F)])

ggplot(dat_res, aes(x = value, y = L1)) + 
    geom_density_ridges(rel_min_height = 0.01) +
    theme_bw() +
    ylab("Item") +
    xlab("Bayesian R2") +
    theme(panel.grid.minor = element_blank())
```

Further customization options can be found [here][https://cran.r-project.org/web/packages/ggridges/vignettes/introduction.html].
